---
title: "Homework 4"
author: "Ericka Smith"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
```

## Problem 1

```{r}
butters <- read_excel("C:/Users/erick/Downloads/butterflies.xlsx")
names(butters)[2] <- "numyears"
```

### Part A

```{r}
summary(glm(LargeSkipper~factor(Region)+numyears, family=poisson, data=butters))
summary(glm(LargeSkipper~factor(Region)*numyears, family=poisson, data=butters))
```

I fit the first model without an interaction term and the second model with an interaction term. For the one without an interaction term all of the coefficients were significant at level $\alpha=0.05$ ($p=0.00294$ for region and $p=0.02796$ for number of years). For the one with an interaction term both the number of years and the region were significant at level $\alpha=0.10$ ($p=0.0581$ for region and $p=0.0507$ for number of years). The AIC is slightly higher for the model with an interaction term, and since the interaction term itself is not significant, I believe the model without an interaction term is the better fit. 

### Part B

#### Pearly Heath

```{r}
summary(glm(PearlyHeath~factor(Region)+ numyears, family=poisson, data=butters))
summary(glm(PearlyHeath~factor(Region)*numyears, family=poisson, data=butters))
```

I fit the first model without an interaction term and the second model with an interaction term. For the one without an interaction term only the coefficient for the number of years was significant at level $\alpha=0.05$ ($p=0.0167$). For the one with an interaction term none of the coefficients were significant. The AIC is slightly higher for the model with an interaction term (137.39 without vs. 135.39 with). I believe the model without an interaction term is clearly the better fit. 

#### Ringlet

```{r}
summary(glm(Ringlet~factor(Region)+ numyears, family=poisson, data=butters))
summary(glm(Ringlet~factor(Region)*numyears, family=poisson, data=butters))
```

I fit the first model without an interaction term and the second model with an interaction term. For the one without an interaction term all of the coefficients were significant at level $\alpha=0.05$ ($p=1.63*10^{-05}$ for region and $p=4.81*10^{-14}$ for number of years). For the one with an interaction term both the number of years and the region were significant at level $\alpha=0.05$  ($p=0.0363$ for region and $p=5.11*10^{-05}$ for number of years), but the interactino term was not significant ($p=0.3618$). The AIC is slightly higher for the model with an interaction term (532.88 compared to 531.71), and since the interaction term itself is not significant, I believe the model without an interaction term is the better fit. 

## Problem 2 (starts next page)


